Methods for understanding the variable importance of local explanations of black-box models

### 2021-11-04 _paper.rmd
starting to make use case figures.

The benefit of having dynamic interaction with data is predicated on a reasonably small render time. It is important to preprocess as much work as possible so that application resources can be used efficiently. Below we discuss the steps and details of the reprocessing.
<!-- Fitting models regularly has non-negligible runtime and the exaction of a single observations local explanations is a similarly computationally expensive operation. -->

@@

<!-- render time -->
The work remaining at runtime consists mostly in rendering the frames of the tour as specified by the selection of the parameters. The application is made with __shiny__ [@chang_shiny_2021]. The global view is created first with __ggplot2__ [@wickham_ggplot2_2016] which is then rendered to an interactive html widget with __plotly__ [@sievert_interactive_2020]. The tour view is created in __spinifex__ [@spyrison_spinifex_2020], which creates the manual tour basis array, and facilitates similar rendering to html widget.

# Software Infrastructure {#sec:softwareinfrastructure}
## Extend spinifex, consume DALEX & treeshap
## Preprocess
## Runtime rendering

@@

>>>>

The benefit of having dynamic interaction with data is predicated on a reasonably small render time. It is important to preprocess as much work as possible so that application resources can be used efficiently. The work remaining at runtime should be solely responding to inputs and the rendering of figures and tables. Below we discuss the steps and details of the reprocessing.

@@

# Software Infrastructure {#sec:softwareinfrastructure}

The above described method and application are implemented as an open source __R__ package, __cheem__ _TODO:XXX site github? cran?_. Preprocessing was facilited with models created via __randomForest__ [liaw_classification_2002], and explanations calculated with __treeshap__ [@kominsarczyk_treeshap_2021].
The application is made with __shiny__ [@chang_shiny_2021]. The tour visual is an extension of __spinifex__ [@spyrison_spinifex_2020]. Both views are created first with first with __ggplot2__ [@wickham_ggplot2_2016] and then rendered as interactive html widget with __plotly__ [@sievert_interactive_2020]. The free ebook, _Explanatory Model Analysis_ [@biecek_explanatory_2021] was a huge boon to understanding local explanations and how to apply them.

Provided with tidy and transformed data, the function `cheem_ls` is a wrapper function that facilitates the creation of random forest models and calculates the tree SHAP for all observations. It returns an a list containing data frames to be consumed by the application, higher level model performance, and runtime information. The return of this function can be then be saved to an .rds file for consumption in the application.

## Linking brushing and highlighting

The dynamic interaction with the global view is critical to the selection of the primary and comparison observations. Linked brushing also for a rectangular selection of observations. These points are highlighted within the other space of the global view _and_ crucially available for downstream consumption, for instance highlighting in the tour. The distinction between the former and latter is subtle, but important for developers to grasp. In the former, self-contained html widgets can highlight directly within javascript without evaluation evaluation downstream in other reactive functions. In the latter, the use of `plotly::event_data()` returns the identity of the selected observations. This will require a reactive flush, but after that, the sky is the ceiling.

@@

We would like to thank Professor Przemyslaw Biecek for his input in early in the project and to the broader MI\^2 lab group for the __DALEX__ ecosystem of __R__ and __Pyhton__ packages. This research was supported in part by Australian government Research Training Program (RTP) scholarships.
